### 2025-05-17 reportdrafting/test.tex
messing around with some latex, early outline

\documentclass[conference]{IEEEtran}
%Template version as of 6/27/2024

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Neural Networks for Neural Tumors}

\author{
\IEEEauthorblockN{Joe Vogel}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jvogel@ucdavis.edu}

\and
\IEEEauthorblockN{Jonathan Mo}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
jtnmo@ucdavis.edu}

\and
\IEEEauthorblockN{Nam Nguyen}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
ncnn@ucdavis.edu}

\and
\IEEEauthorblockN{Aryan Karnwal}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
akarnwal@ucdavis.edu}

\and
\IEEEauthorblockN{Richard Ho}
\IEEEauthorblockA{\textit{UC Davis Computer Science} \\
rdho@ucdavis.edu}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert.
\end{IEEEkeywords}

\section{Introduction}
This document is a model and instructions for \LaTeX.
Please observe the conference page limits. For more information about how to become an IEEE Conference author or how to write your paper, please visit   IEEE Conference Author Center website: https://conferences.ieeeauthorcenter.ieee.org/.

\end{document}